MATH 245 Homework 2

### 2024-02-16 PDEs/math245_hw2.Rmd
finished PDE prob 8

---
title: "MATH 245 Homework 2"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-15"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

@@

## Done, feel OK: 1, 2, 5, 7
## Done, very low confidence: 4
## Redo: 6
## Not started: 3, 8

@@

## Question 2
Using a factorization similar to the wave equation, solve the following IVP:
\begin{equation}
    \begin{cases}
      u_{xx}+2u_{xy}-3u_{yy}=0 & x\subset \mathbb{R},\; y> 0\\
      u(0,x)=\sin{x} & x\subset \mathbb{R}\\
      u_y(0,x) = x & x\subset \mathbb{R}
    \end{cases}       
\end{equation}

@@

\begin{equation}
\begin{cases}
u_x-u_y=f(3x-y)
u_x+3u_y=g(x+y)
\end{cases}
\end{equation}

@@

Solve the Neumann boundary value problem for the wave equation on half line:
\begin{equation}
    \begin{cases}
      u_{tt}=c^2u_{xx}+f(t,x) & 0<x<\infty\\
      u(0,x)=\phi{x} & 0<x<\infty\\
      u_t(0,x) = \psi{x} & 0<x<\infty\\
      u_x(t,0) = h(t) & t>0
    \end{cases}       
\end{equation}

@@

\begin{equation} u_t(0,r) = \begin{cases}
  1  & \lvert{r}\rvert \leq 1 \\
  0 &  \lvert{r}\rvert > 1
\end{cases}
\end{equation}

@@

Since $u=\frac{w}{r}$, this means we have
\begin{tcolorbox}[colback=white, title=Solution]
\begin{equation}u(t,r)=
    \begin{cases} 
      t & \text{ if } r-ct \geq -1, r+ct \leq 1\\
      0 & \text{ if } r-ct < -1, r+ct > 1\\
      \frac{1}{4rc} ((r+ct)^2-1) & \text{ if } r-ct < -1, r+ct \leq 1\\
      \frac{1}{4rc} (1-(r-ct)^2) &\text{ if }  r-ct \geq -1, r+ct > 1\\
    \end{cases}       
\end{equation}
\end{tcolorbox}

## Question 5
Consider the following Dirichlet boundary value problem:
\begin{equation} 
\begin{cases}
  u_{tt}+x(t,x)u_t=u_{xx} & 0<x<1 \\
  u(0,x)=\phi(x) & 0<x<1 \\
  u_t(0,x)=\psi(x) & 0<x<1 \\
  u(t,0) = u(t,1)=0 & t \geq 0 \\
\end{cases}
\end{equation}

@@

(4) Use the previous part to prove uniqueness of the following problem:
\begin{equation} 
\begin{cases}
  u_{tt}+a(t,x)u_t=u_{xx} & 0<x<1, t>0 \\
  u(0,x)=\phi(x) & 0<x<1 \\
  u_t(0,x)=\psi(x) & 0<x<1 \\
  u(t,0) = f(t) & t \geq 0 \\
  u(t,1) = g(t) & t \geq 0 \\
\end{cases}
\end{equation}

@@

\begin{equation} 
\begin{cases}
  u_{tt}-u_{xx}- \frac{2}{x}u_x=0 & -\infty <x<\infty, t>0 \\
  u(0,x)=0 & -\infty <x<\infty \\
  u_t(0,x)=g(x) & -\infty <x< \infty \\
\end{cases}
\end{equation}

@@

## Problem 8   
Solve the following characteristic initial value problem:
\begin{equation} 
\begin{cases}
  y^3u_{xx}-yu_{yy}+u_y=0 & 0<x<4, \quad \lvert y \rvert \leq 2 \sqrt{2} \\
  u(x,y)= f(x) & x+ \frac{y^2}{2}=4 \text{ for } 2 \leq x \leq 4 \\
  u(x,y)=g(x) & x- \frac{y^2}{2}=0 \text{ for } 0 \leq x \leq 2 \\
\end{cases}
\end{equation}

@@

When $x=2$,

$f(x)=h_1(4)+h_2(2x-4)$

$z=2x-4, \quad x=\frac{z}{2} + 2$

$h_2(z)=f\left(\frac{z}{2}+2\right)-h_1(4)$

and
$g(x)=h_1(2x)+h_2(0)$

$h_1(z)=g\left(\frac{z}{2}\right)-h_2(0)$

$h_1+h_2=g\left(\frac{z}{2}\right)-h_2(0)+f(\frac{z}{2} + 2)-h_1(4)$

@@

so after replacing $z$ by $x+\frac{y^2}{2}$ and $x-\frac{y^2}{2}$, we get

\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=g\left(\frac{x}{2}+\frac{y^2}{4}\right)+f\left(\frac{x}{2}+\frac{y^2}{4}+2\right)+f(2)$$
\end{tcolorbox}

>>>>

---
title: "MATH 245 Homework 2"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-16"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

@@

## Done, feel OK: 1, 2, 5, 7, 8
## Done, low confidence: 4
## Redo: 6
## Not started: 3

@@

## Question 2
Using a factorization similar to the wave equation, solve the following IVP:
\begin{equation*}
    \begin{cases}
      u_{xx}+2u_{xy}-3u_{yy}=0 & x\subset \mathbb{R},\; y> 0\\
      u(0,x)=\sin{x} & x\subset \mathbb{R}\\
      u_y(0,x) = x & x\subset \mathbb{R}
    \end{cases}       
\end{equation*}

@@

\begin{equation*}
\begin{cases}
u_x-u_y=f(3x-y)
u_x+3u_y=g(x+y)
\end{cases}
\end{equation*}

@@

Solve the Neumann boundary value problem for the wave equation on half line:
\begin{equation*}
    \begin{cases}
      u_{tt}=c^2u_{xx}+f(t,x) & 0<x<\infty\\
      u(0,x)=\phi{x} & 0<x<\infty\\
      u_t(0,x) = \psi{x} & 0<x<\infty\\
      u_x(t,0) = h(t) & t>0
    \end{cases}       
\end{equation*}

@@

\begin{equation*} u_t(0,r) = \begin{cases}
  1  & \lvert{r}\rvert \leq 1 \\
  0 &  \lvert{r}\rvert > 1
\end{cases}
\end{equation*}

@@

Since $u=\frac{w}{r}$, this means we have
\begin{tcolorbox}[colback=white, title=Solution]
\begin{equation*}u(t,r)=
    \begin{cases} 
      t & \text{ if } r-ct \geq -1, r+ct \leq 1\\
      0 & \text{ if } r-ct < -1, r+ct > 1\\
      \frac{1}{4rc} ((r+ct)^2-1) & \text{ if } r-ct < -1, r+ct \leq 1\\
      \frac{1}{4rc} (1-(r-ct)^2) &\text{ if }  r-ct \geq -1, r+ct > 1\\
    \end{cases}       
\end{equation*}
\end{tcolorbox}

## Question 5
Consider the following Dirichlet boundary value problem:
\begin{equation*} 
\begin{cases}
  u_{tt}+x(t,x)u_t=u_{xx} & 0<x<1 \\
  u(0,x)=\phi(x) & 0<x<1 \\
  u_t(0,x)=\psi(x) & 0<x<1 \\
  u(t,0) = u(t,1)=0 & t \geq 0 \\
\end{cases}
\end{equation*}

@@

(4) Use the previous part to prove uniqueness of the following problem:
\begin{equation*} 
\begin{cases}
  u_{tt}+a(t,x)u_t=u_{xx} & 0<x<1, t>0 \\
  u(0,x)=\phi(x) & 0<x<1 \\
  u_t(0,x)=\psi(x) & 0<x<1 \\
  u(t,0) = f(t) & t \geq 0 \\
  u(t,1) = g(t) & t \geq 0 \\
\end{cases}
\end{equation*}

@@

\begin{equation*} 
\begin{cases}
  u_{tt}-u_{xx}- \frac{2}{x}u_x=0 & -\infty <x<\infty, t>0 \\
  u(0,x)=0 & -\infty <x<\infty \\
  u_t(0,x)=g(x) & -\infty <x< \infty \\
\end{cases}
\end{equation*}

@@

## Problem 8   
Solve the following characteristic initial value problem:
\begin{equation*} 
\begin{cases}
  y^3u_{xx}-yu_{yy}+u_y=0 & 0<x<4, \quad \lvert y \rvert \leq 2 \sqrt{2} \\
  u(x,y)= f(x) & x+ \frac{y^2}{2}=4 \text{ for } 2 \leq x \leq 4 \\
  u(x,y)=g(x) & x- \frac{y^2}{2}=0 \text{ for } 0 \leq x \leq 2 \\
\end{cases}
\end{equation*}

@@

Now we use our given conditions. When  $2 \leq x \leq 4$, we have $g(x)=h_1(2x)+h_2(0)$, which means $h_1(z_1)=g\left(\frac{z_1}{2}\right)-h_2(0)$.

Similarly, when $0 \leq x \leq 2$, we have $f(x)=h_1(4)+h_2(2x-4)$, which means $h_2(z_2)=f\left(\frac{z_2}{2}+2\right)-h_1(4)$.

Thus,
$$h_1(z_1)+h_2(z_2)=g\left(\frac{z_1}{2}\right)-h_2(0)+f\left(\frac{z_2}{2} + 2\right)-h_1(4)$$

@@

so after replacing $z_1$ by $x+\frac{y^2}{2}$ and $z_2$ by $x-\frac{y^2}{2}$, we get

\begin{tcolorbox}[colback=white, title=Solution]
$$u(x,y)=g\left(\frac{x}{2}+\frac{y^2}{4}\right)+f\left(\frac{x}{2}+\frac{y^2}{4}+2\right)-f(2)$$
\end{tcolorbox}

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Check solution]
$$u_y(x,y)=\frac{y}{2}g'\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{y}{2}f'\left(\frac{x}{2}+\frac{y^2}{4}+2\right)$$

$$u_{yy}(x,y)=\frac{y^2}{4}g''\left(\frac{x}{2}+\frac{y^2}{4}\right)+
\frac{1}{2}g'\left(\frac{x}{2}+\frac{y^2}{4}\right)+
\frac{y^2}{4}f''\left(\frac{x}{2}+\frac{y^2}{4}+2\right)+
\frac{1}{2}f'\left(\frac{x}{2}+\frac{y^2}{4}+2\right)$$

$$u_x(x,y)=\frac{1}{2}g'\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{2}f'\left(\frac{x}{2}+\frac{y^2}{4}+2\right)-f'(2)$$

$$u_{xx}(x,y)=\frac{1}{4}g''\left(\frac{x}{2}+\frac{y^2}{4}\right)+\frac{1}{4}f''\left(\frac{x}{2}+\frac{y^2}{4}+2\right)-f''(2)$$

\begin{equation*}
  \begin{aligned}
  & \left[\frac{y^3}{4} g''\left(\frac{x}{2} +\frac{y^2}{4}\right) + \frac{y^3}{4} f''\left(\frac{x}{2}+\frac{y^2}{4}+2\right)\right]- \\
  & \left[\frac{y^3}{4}g''\left(\frac{x}{2}+\frac{y^2}{4}\right)+
\frac{y}{2} g' \left(\frac{x}{2}+\frac{y^2}{4} \right)+
\frac{y^3}{4} f'' \left( \frac{x}{2}+ \frac{y^2}{4}+2 \right)+
\frac{y}{2}f' \left( \frac{x}{2}+ \frac{y^2}{4}+2 \right) \right] + \\
  & \left[ \frac{y}{2} g' \left( \frac{x}{2}+ \frac{y^2}{4} \right)+ \frac{y}{2}f' \left( \frac{x}{2}+ \frac{y^2}{4}+2 \right) \right]=0
 \end{aligned}
\end{equation*}
